---
title: "Transmission and  Mile Per Gallon (MPG)"
author: "Cynthia Tang"
date: "June 9, 2019"
output:
  md_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Summary  
The report explored the relationship between transmission types and miles per gallon (MPG). Figure 1 in appendix showed that an manual transmission trend to be better for MPG, considering other factors. Then, we performed t-test to verify the hypothesis. To quantify the MPG difference between automatic and manual transmissins, we fitted the MPG (outcome) and a set of variables (regressors) in to multiple linar models. We selected tansmission (am), number of cylinders (cyl), gross horsepower (hp), and weight (wt) as repressors.  

## Results
```{r, include=FALSE}
mtcars1 <- mtcars
mtcars1$am <- ifelse(mtcars$am == 0, "automatic", "manual")

# t.test
t.am <- t.test(mpg ~ factor(am), data = mtcars1, alternative = "less")
# fitting lm
library(car)
fit1 <- lm(mpg ~ factor(am), data =mtcars1)
fit2 <- lm(mpg ~ factor(am) + factor(cyl), data =mtcars1)
anova(fit1, fit2)
vif(fit2)
fit3 <- lm(mpg ~ factor(am) + factor(cyl) + disp, data =mtcars1)
vif(fit3)
anova(fit1, fit2, fit3) # it unnessary to include disp as a regressor
fit4 <- lm(mpg ~ factor(am) + factor(cyl) + hp, data =mtcars1)
vif(fit4)
anova(fit1, fit2, fit4) # hp increased variance
fit5 <- lm(mpg ~ factor(am) + factor(cyl) + hp + drat, data =mtcars1)
vif(fit5)
anova(fit1, fit2, fit4, fit5) # unnessary
fit6 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt, data =mtcars1) # selected!
vif(fit6)
anova(fit1, fit2, fit4, fit6) # wt increased variance
fit7 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + qsec, data =mtcars1)
vif(fit7)
anova(fit1, fit2, fit4, fit6, fit7) # unnecessary
fit8 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + factor(vs), data =mtcars1)
vif(fit8)
anova(fit1, fit2, fit4, fit6, fit8) # unnecessary
fit9 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + factor(gear), data =mtcars1)
vif(fit9)
anova(fit1, fit2, fit4, fit6, fit9) # unnecessary
fit10 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + factor(carb), data =mtcars1)
vif(fit10)
anova(fit1, fit2, fit4, fit6, fit10) # unnecessary
summary(fit6)
```

The mean of automatic transmission: `r t.am$estimate[1]` miles per gallon.  
The mean of manual transmission: `r t.am$estimate[2]` miles per gallon.   
  
__Model Selection__  
We used nested model search. The detailed process of model selection is shown in __Appendix__.  
We set mpg as the outcome and transmission (am), number of cylinders (cyl), gross horsepower (hp), and weight (wt) as repressors and fitted the linear model: $MPG_i = \beta_0 + \beta_1 am(manual) + \beta_2 cyl(6) + \beta_3 cyl(8) + \beta_4 hp_i + \beta_5 wt_i + \epsilon_i$. Assume $\epsilon_i$ ~ $N(\mu, \sigma^2)$.     
The coefficients of the linear model are showed below:
```{r}
coefficients(fit6)
```

__Interpretion__   
The intercept $\beta_0$: the expected mpg of automatic transmission when there is 4 cylinders, 0 gross horsepower, and none weight.  
The $\beta_1$: the expected change in mpg comparing those in manual to those in automatic transmission.  
The $\beta_2$: the expected change in mpg comparing those with 6 cylinders to those with 4 cylinders.  
The $\beta_3$: the expected change in mpg comparing those with 8 cylinders to those with 4 cylinders.  
The $\beta_4$: the expected change in mpg for every 1 gross horsepower increase.  
The $\beta_5$: the expected change in mpg fot every 1000 lbs increase in the weight of cars.  
  
__T-test__  
Assumption: Unequal variance. Detailed test results is shown in __Appendix__.     
$H_0$: $\mu_{auto} = \mu_{manual}$.  
$H_a$: $\mu_{auto} < \mu_{manual}$.  
The p-value: `r t.am$p.value`. The type I error $\alpha$ = 0.05, then we rejected $H_0$ and concluded that a manual transmission is better for MPG.  
  
__Quantify the MPG Difference__  
$\mu_{manual} - \mu_{auto}$ = `r t.am$estimate[2] - t.am$estimate[1]`.  
Adjustment: $MPG_{manual} - MPG_{auto}$ = abs($\beta_1$) = `r abs(coefficients(fit6)[2])`. 
  
__Diagnostics__  
```{r diagnostic, fig.width=5, fig.height=4}
par(mfrow = c(2, 2))
plot(fit6)
```
  
The mean of residuals is `r mean(fit6$residuals)`, approximating to zero.   
"Residual vs Fitted": residuals are independent. "scale-location": the homoscedasticity of residuals. "Normal Q-Q"": all residuals were approximately stardard normal distributed. "Residual vs Leverge": within 0.5, no outliers.
  
__The uncertainty__   
We assumed other variables not included in the linear model were completely randomized. The error $\epsilon$ in the linear model followed normal distribution.
However, we are uncertain whether there is a better linear model with other combination of diferent regressors.   

## Appendix  
### Figure  
```{r plots, message=FALSE, warning=FALSE, fig.width=7, fig.height=5}
library(ggplot2)
library(gridExtra)
# plotting
p1 <- ggplot(mtcars1, aes(am, mpg, fill = am)) +
        geom_boxplot() +
        xlab("Transmission") +
        ylab("mpg (miles/(US) gallon") +
        ggtitle("MPG vs Transmission")
p2 <- ggplot(mtcars1, aes(wt, mpg, color = am)) +
        geom_point() +
        xlab("Weight (1000lbs)") +
        ylab("mpg (miles/(US) gallon") +
        ggtitle("MPG vs Transmission and Weight")
p3 <- ggplot(mtcars1, aes(hp, mpg, color = am)) +
        geom_point() +
        xlab("Gross horsepower") +
        ylab("mpg (miles/(US) gallon") +
        ggtitle("MPG vs Transmission and \n Gross horsepower")
p4 <- ggplot(mtcars1, aes(factor(cyl), mpg)) +
        geom_point(aes(color = am)) +
        geom_boxplot(aes(fill = am), alpha = 1/5) +
        xlab("Number of Cylinders") +
        ylab("mpg (miles/(US) gallon") +
        ggtitle("MPG vs Transmission and \n the Number of Cylinders")
grid.arrange(p1, p2, p3, p4, bottom = "Figure 1. The relationship between MPG and a set of variables", nrow = 2)
```
  
### Model Selection  
```{r modelselection}
# fitting lm
library(car)
fit1 <- lm(mpg ~ factor(am), data =mtcars1)
fit2 <- lm(mpg ~ factor(am) + factor(cyl), data =mtcars1)
anova(fit1, fit2)
fit3 <- lm(mpg ~ factor(am) + factor(cyl) + disp, data =mtcars1)
anova(fit1, fit2, fit3) # it unnessary to include disp as a regressor
fit4 <- lm(mpg ~ factor(am) + factor(cyl) + hp, data =mtcars1)
anova(fit1, fit2, fit4) # hp increased variance
fit5 <- lm(mpg ~ factor(am) + factor(cyl) + hp + drat, data =mtcars1)
anova(fit1, fit2, fit4, fit5) # unnessary
fit6 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt, data =mtcars1) # selected!
anova(fit1, fit2, fit4, fit6) # wt increased variance
fit7 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + qsec, data =mtcars1)
anova(fit1, fit2, fit4, fit6, fit7) # unnecessary
fit8 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + factor(vs), data =mtcars1)
anova(fit1, fit2, fit4, fit6, fit8) # unnecessary
fit9 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + factor(gear), data =mtcars1)
anova(fit1, fit2, fit4, fit6, fit9) # unnecessary
fit10 <- lm(mpg ~ factor(am) + factor(cyl) + hp + wt + factor(carb), data =mtcars1)
anova(fit1, fit2, fit4, fit6, fit10) # unnecessary
```

### T-test Results  
```{r ttest}
# t.test
t.test(mpg ~ factor(am), data = mtcars1, alternative = "less")

```

```{r peers, include=FALSE}
fullModelFit <- lm(mpg ~ ., data = mtcars)
stepFit <- step(fullModelFit)
summary(stepFit) # results hidden
summary(stepFit)$coeff # results hidden
sum((abs(dfbetas(stepFit)))>1)
```

